Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-11-18 ms.tex
Finishing up Phil's last edits up to discussion section

The emissivity for an impulsive source with energy E, source position $\vec{x}_s$, and frequency $\nu_s$ is derived in Appendix \ref{app:rteqn_derivation}. Considering a photon source at $\vec{x}_s=0$, we have (Eq. \ref{eq:jem})
\be
j_{\rm em} & = & \frac{E}{4\pi} \delta^3(\vec{x}) \delta(\nu-\nu_{\rm s})\delta (t) ,
\label{eq:jem2}
\ee 
The resulting equation for $J(r,\sigma,t)$ is
\be
-3 \frac{k\phi}{c} \frac{\partial J}{\partial t} + \nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2}
& = & - \frac{\sqrt{6} kE}{4\pi \Delta^2} \delta^3(\vec{x}) \delta (\sigma - \sigma_s ) \delta (t).
\label{eq:diffusion_eqn}
\ee
We employ an expansion in terms of spherical Bessel functions in $r$ and Fourier transform in time. The expansion for $J(r, \sigma, t)$ is then
\be
\label{eq:jrsigmat_expansion}
J(r, \sigma, t) = \sum_{n=1}^{\infty} \int_{-\infty}^\infty \frac{d\omega}{2\pi} e^{-i\omega t} j_0\left(\kappa_n r\right) J(n, \sigma, \omega),
\ee
with
\be \label{eq:jnsigmaomega}
J(n, \sigma, \omega) = \frac{2\kappa_n^2}{R} \int_0^R dr\ r^2 j_0(\kappa_n r) \int_{-\infty}^\infty dt\ e^{i\omega t} J(r, \sigma, t).
\ee
Here, $\omega$ describes the time-dependence of $J$. Though it is written as a function of the photon frequency variable $\sigma$, $J(r, \sigma, t)$ is the specific mean intensity $dE/(dA dt d\nu)$ and is a distribution in $\nu$. The Fourier coefficient $J(n, \sigma, \omega)$ has units $dE/(dA d\nu)$. Using Equation \ref{eq:jnsigmaomega}, we obtain
\be \label{eq:diffusion_plugged_in}
 \left( \frac{3k\phi}{c}i\omega  -   \kappa_n^2 \right) J(n,\sigma,\omega)  &+& \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2J(n,\sigma,\omega)}{\partial\sigma^2} = -\frac{2\kappa_n^2}{R} \frac{\sqrt{6}}{4\pi} \frac{kE}{\Delta^2} \frac{1}{4\pi} \delta(\sigma - \sigma_s).
\ee
At $\sigma=\sigma_s$, continuity must be enforced,
\be \label{eq:matching_condition_1}
J(n, \sigma^-, \omega) = J(n, \sigma^+, \omega),
\ee
and the discontinuity in $dJ/d\sigma$ due to the source is
\be \label{eq:matching_condition_2}
\frac{\partial J(n, \sigma^+, \omega)}{\partial \sigma} - \frac{\partial J(n, \sigma^-, \omega)}{\partial \sigma} & = & 
- \frac{\sqrt{6}}{8} n^2 \frac{E}{kR^3}.
\ee
At large values of $\sigma$ the line profile $\phi$ is small and Eq. \ref{eq:diffusion_plugged_in} becomes
\be \label{eq:diffusion_at_large_sigma}
\frac{\partial^2J}{\partial\sigma^2} \approx \frac{\Delta^2\kappa_n^2}{k^2} J,
\ee
which has solutions 
\be
J(n, \sigma, \omega)\ {\sim}\ e^{\pm \Delta \kappa_n \sigma / k}.
\ee
This approximate solution implies a boundary condition at large $|\sigma|$
\be \label{eq:single_j_derivative}
\frac{\partial J}{\partial \sigma} = \mp \frac{\Delta\kappa_n}{k} J,
\ee
where a negative sign is taken for large $+\sigma$ and a positive sign is taken for large $-\sigma$ to chose the finite solution as $|\sigma|\to 0$. Numerical integrations are performed inward toward $\sigma_s$ over several domains: from large $|\sigma|$ to $\sigma_s$, from large $|\sigma|$ to 0, and from 0 to $\sigma_s$, depending on whether $\sigma_s$ is positive or negative. If $\sigma_s=0$, just two integrations are performed inward from large $|\sigma|$ to 0. Initial values for integration are obtained either by setting $J=1$ and $dJ/d\sigma$ from Equation \ref{eq:single_j_derivative} at large $|\sigma|$ or by matching $J$ and $dJ/d\sigma$ at 0. This gives $J$ and $J'$ on either side of $\sigma_s$, where a prime indicates the derivative $\partial/\partial \sigma$. By enforcing the matching conditions Eqs. \ref{eq:matching_condition_1} and \ref{eq:matching_condition_2}, the eigenfunctions $J(n, \sigma, \omega)$ are obtained over the domain of photon frequencies $\sigma$. Since the solutions are linear in the starting conditions, only two integrations with different starting values are necessary.

We now employ an ansatz that roughly agrees with Monte Carlo results for $x_s=0$ based on numerical calculations of this result. However, the analysis presented in Appendix \ref{app:wkb} suggests the ansatz is incomplete in the case where $x_s \neq 0$ and the solution is asymmetric about the line center. Let us define the damping rate to be $\gamma \equiv i\omega$, which is real and positive for damped solutions. At the eigenvalues $\gamma=\gamma_{nm}$, the response $J(n, \sigma, \omega)$ is resonant. Near these $\gamma_{nm}$ poles, the eigenfunctions will have the form 
\be \label{eq:jnsigmaomega_approx}
J(n,\sigma,-i\gamma) & \simeq \frac{ J_{nm}(\sigma) }{\gamma_{nm} - \gamma} + C(\sigma),
\ee
where $C(\sigma)$ varies slowly in $\gamma$. An application of the residue theorem gives
\be
J(r,\sigma,t) & \simeq & j_0(\kappa_n r) J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Summing over all spatial modes $n$ and over all eigenmodes $m$ for a given $n$, we obtain
\be \label{eq:Jrsigmat}
J(r,\sigma,t) & = & \sum_{n=1}^\infty j_0(\kappa_n r)  \sum_{m=1}^{\infty} J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
This ansatz captures the contributions from $n \times m$ simple poles. Taking a derivative with respect to $r$ and evaluating at the surface $r=R$, we use
\be
\frac{dj_0(\kappa_n r)}{dr} \bigg\rvert_R & =& \frac{d}{dr} \left[ \frac{\sin(\kappa_n r)}{\kappa_n r} \right]\bigg\rvert_R
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)
\nonumber \\ & = & \frac{(-1)^n}{R}
\ee
to obtain the flux, which is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Multiplying by $4\pi R^2$ gives the energy per time per frequency emerging from the sphere to be
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\label{eq:dEdtdnu}
\ee
Integrating over time yields a factor $1/\gamma_{nm}$, and by integrating over $d\nu$ we find
\be \label{eq:sum_rule}
1 & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE} \sum_{nm} (-1)^{n+1} \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial ``sum rule'' provides a check on the values of $\gamma_{nm}$ and $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each mode is
\be \label{eq:pnmsoln}
P_{nm} & \equiv & \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^{n+1} \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients are negative for odd values of $n$ and positive for even $n$. The size of each contribution scales roughly as $0.5/(m-7/8)^{2/3}$, with a weak dependence on $n$. This indicates the need for a large number of $n$ and $m$ to converge, in that it takes roughly ten times as many $m$ modes to reduce the size of $P_{nm}$ by a factor of ${\sim}$5.

@@

We seek now to calculate the eigenmodes $J_{nm}(\sigma)$ and eigenfrequencies $\gamma_{nm}$ for a given spatial $n$. These will be labelled by an index $m=1, 2, ...$. To measure the size of the response to detect where resonances occur, we sum the absolute value of $J(n,\sigma,-i\gamma)$ over the array $\sigma$. We call this response $f$, and use the index $j$ to represent the value of the response at discrete points $\gamma_j$ over a range of $\gamma$. In places where $f_j > f_{j-1}$ and $f_j>f_{j+1}$, we have bracketed a resonance that occurs in the interval $(\gamma_{j-1},\gamma_{j+1})$. We then hone in on the eigenvalue before continuing the sweep in $\gamma$. To refine the eigenmode solution, we evaluate $f_{j-1}$, $f_j$, and $f_{j+1}$ at the points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$. Assuming the form in Equation \ref{eq:jnsigmaomega_approx}, a guess at the correct eigenvalue $\gamma_{nm}$ can be calculated from
\be
\gamma_{\rm guess} &=& \frac{b\gamma_{j-1} - \gamma_{j+1}}{b - 1},
\ee
where
\be
b &=& \left(\frac{f_{j} - f_{j-1}}{f_{j} - f_{j+1}}\right)\left(\frac{\gamma_{j+1}-\gamma_{j}}{\gamma_{j-1}-\gamma_{j}}\right).
\ee
This is repeated by replacing initial points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$ with closer estimates while the size of the response grows as the resonance is approached. After iterating an eigenvalue $\gamma_{nm}$ to convergence, we now find the corresponding eigenfunction $J_{nm}(\sigma)$. We can solve for $J_{nm}(\sigma)$ by comparing two points $\gamma_1$ and $\gamma_2$ near the resonance, cancelling $C(\sigma)$ in the difference to find
\be
J_{nm}(\sigma) & \simeq & \frac{ J(n,\sigma,-i\gamma_1)  - J(n,\sigma,-i\gamma_2) }{ 1/(\gamma_{nm}-\gamma_1) - 1/(\gamma_{nm}-\gamma_2)}.
\ee
\begin{figure}
    \centering
    \includegraphics{Jsoln_n1_m100.pdf}
    \caption{Eigenfunctions $J_{nm}(x)$ for the lowest-order spatial eigenmode $n=1$, and $m=1, ..., 100$ with $x_s=0$, $a = 4.72\times 10^{-4}$ and $\tau_0=10^7$.}
    \label{fig:jsoln}
\end{figure}
\noindent The form of a single eigenmode $J_{nm}(\sigma)$ is oscillatory out to some turning point, $\sigma_{tp}$, at which point the function becomes evanescent. The location of the turning point can be found by ignoring the delta-function discontinuity at the source frequency $\sigma_s$ in Eq. \ref{eq:diffusion_plugged_in} and examining the resulting homogeneous differential equation. We obtain
\be \label{eq:wkb_differential_eqn}
\frac{d^2J}{d\sigma^2} & = & \left[ \left( \frac{\kappa_n \Delta }{k} \right)^2 - \frac{3\phi \gamma\Delta^2}{ck}\right] J,
\ee
where the line profile is approximated as in Eq. \ref{eq:app:line_profile_wing}. When the coefficient on the right hand side is positive, exponential growth or decaying evanescent solutions are found. This occurs on the line wings. When the coefficient on the right hand side is negative, oscillatory solutions are found (propagation), which occurs near the line core. The boundary between propagation and evanescence occurs at the turning point, given by
\be \label{eq:sigma_tp}
\sigma_{\rm tp} & = & \sqrt{\frac{2a}{\pi}}\left( \frac{k \gamma}{ \kappa_n^2 c \Delta} \right)^{3/2}
\ee
Thus, to ensure accuracy in each term of Eq. \ref{eq:Jrsigmat}, the bounds of $\sigma$ must be set sufficiently far outside of $\sigma_{\rm tp}$ such that the function is small at the edges. The scale of an $e$-folding in $J_{nm}(\sigma)$ is $k/(\kappa_n \Delta) = \tau_0 / (\sqrt{\pi} n)$, so a grid of $\sigma$ is chosen that spans a large enough number of $e$-foldings that no oscillatory behavior is present at the boundaries of the calculation.

The eigenfunction's oscillatory forms have varying amplitudes which sum to create the final form of the mean intensity. The largest contribution always comes from the ($n=1, m=1$) lowest-order eigenfunction. Fig. \ref{fig:jsoln} shows a set of eigenfunctions $J_{nm}(\sigma)$ to illustrate their relative scales for different $m$ at a fixed spatial eigenmode $n$. The overall scale of the $J_{nm}(\sigma)$ are set by the factor $E/(kR^3)$ with $E$ arbitrarily set to 1, $k = \tau_0 \sqrt{\pi} \Delta / R$, and $R = 10^{11}$ cm. For H atoms with $T=10^4$ K and $\tau_0=10^7$, an eigenfunction has typical size ${\sim} E a / \left(R^2 \Delta \right) = 10^{-37}$. Additional terms add smaller-magnitude, faster-oscillating components that lead to higher accuracy upon summation with the lower-order terms. The oscillations of various modes must cancel, so many modes $m$ and $n$ are required for convergence to the correct solution. 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{gamma_nm.pdf}
    \caption{The scaling of numerically-obtained resonant frequencies as compared with the analytic expression in Eq. \ref{eq:gamma_nm}. The data shown is for $\tau_0=10^7$ and $x_s=0.0$.}
    \label{fig:gamma_nm}
\end{figure}
The values of the $\gamma_{nm}$ can be described approximately with Eq. \ref{eq:gamma_nm}. Their values depend on $m$, $n$, and other physical parameters according to 
\be \label{eq:gamma_nm}
\gamma_{nm} = 2^{-1/3} \pi^{13/6} n^{4/3}\left(m-\frac{7}{8}\right)^{2/3}\frac{c}{R}(a\tau_0)^{-1/3}
\ee
as shown in Appendix \ref{app:wkb}. The power law in $m$ is weak, requiring up to $m=500$ to reduce the scale of $\gamma_{nm}^{\ \ -1}$ by two orders of magnitude. When sweeping through to find resonances, Equation \ref{eq:gamma_nm} is used to set the scale of the sweep points $\gamma_j$ to ensure no $\gamma_{nm}$ are missed. The continuity and close agreement with the analytic expression shown in Fig. \ref{fig:gamma_nm} indicate that the numerical solutions are accurate.

@@

We now calculate the distribution of the times it takes each photon to escape the sphere. This is obtained by integrating Eq. \ref{eq:dEdtdnu} over all frequencies. We find
\be
P(t)  & = & \sqrt{\frac{3}{2}} \frac{16\pi^2 R \Delta^2 }{3kE}     \sum_{nm} (-1)^n  e^{-\gamma_{nm}t} \int d\sigma J_{nm}(\sigma) 
\nonumber \\ & = &  \sum_{nm} P_{nm} \gamma_{nm} e^{-\gamma_{nm}t},
\label{eq:waittime}
\ee
which is normalized to unity. For a sufficiently large number of spatial modes $n$ and frequency modes $m$, the result of this sum can agree with Monte Carlo escape time distributions when $x_s=0$, but it is not a fast enough computational method to use as a distribution for sampling. However, the late-time distribution is simply an exponential falloff. The rate constant of the falloff is the lowest-order eigenfrequency, $\gamma_{11}$, and its scale is determined by the coefficient $P_{11}$ as in Eq. \ref{eq:pnmsoln}. Thus, an approximate ``fitting function'' that captures both the peak of the escape time distribution and the exponential falloff is
\be \label{eq:fitting_function}
P(t) = \exp{\left[-\left(\frac{t_{\rm diff}}{t}\right)^2\right]} \times \gamma_{11} P_{11} e^{-\gamma_{11}t}.
\ee
The first term represents the early-time distribution, which then transitions to an exponential falloff past a point $ct_{\rm diff}/R = (a\tau_0)^{1/3}$, where $t_{\rm diff}$ is the characteristic diffusion timescale.

In Fig. \ref{fig:tau_scaling}, it is shown that the time constant of exponential decay in fitted Monte Carlo escape time distributions converges with $\gamma_{11}^{-1}$ at sufficiently high $\tau_0$, following a $t\propto(a\tau_0)^{1/3}$ scaling. At lower $\tau_0$, the effects of line core scattering are most important, leading to a larger discrepancy in the characteristic escape timescale. Here, the Monte Carlo accurately characterizes the photons which scatter in the core many times before escaping, while the semi-analytic solution does not capture this behavior as it uses only the Lorentzian piece of the line profile and does not use enough spatial modes to accurately model the frequency regime near line center. However, as $\tau_0$ grows, the effect of core scattering becomes smaller and the approximations hold, agreeing well with the expected $(a\tau_0)^{1/3}$ scaling for the rate constant.

>>>>

The emissivity for an impulsive source with energy E, source position $\vec{x}_s$, and frequency $\nu_s$ is derived in Appendix \ref{app:rteqn_derivation}. Considering a photon source at $\vec{x}_s=0$, we have (Eq. \ref{eq:jem})
\be
j_{\rm em} & = & \frac{E}{4\pi} \delta^3(\vec{x}) \delta(\nu-\nu_{\rm s})\delta (t) ,
\label{eq:jem2}
\ee 
The resulting equation for $J(r,\sigma,t)$ is
\be
 \frac{-3k\phi}{c} \frac{\partial J}{\partial t} + \nabla^2 J + \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2 J}{\partial \sigma^2}
& = & - \frac{\sqrt{6} kE}{4\pi \Delta^2} \delta^3(\vec{x}) \delta (\sigma - \sigma_s ) \delta (t).
\label{eq:diffusion_eqn}
\ee
We employ an expansion in terms of spherical Bessel functions in $r$ and Fourier transform in time. The zeroth spherical Bessel function is $j_0(x) = \sin{x}/x$. The expansion for $J(r, \sigma, t)$ is then
\be
\label{eq:jrsigmat_expansion}
J(r, \sigma, t) = \sum_{n=1}^{\infty} \int_{-\infty}^\infty \frac{d\omega}{2\pi} e^{-i\omega t} j_0\left(\kappa_n r\right) J(n, \sigma, \omega),
\ee
with
\be \label{eq:jnsigmaomega}
J(n, \sigma, \omega) = \frac{2\kappa_n^2}{R} \int_0^R dr\ r^2 j_0(\kappa_n r) \int_{-\infty}^\infty dt\ e^{i\omega t} J(r, \sigma, t).
\ee
Here, $\kappa_n = n\pi/R$ and $\omega$ describes the time-dependence of $J$. Though it is written as a function of the photon frequency variable $\sigma$, $J(r, \sigma, t)$ is the specific mean intensity $dE/(dA dt d\nu)$ and is a distribution in $\nu$. The Fourier coefficient $J(n, \sigma, \omega)$ has units $dE/(dA d\nu)$. Using Equation \ref{eq:jnsigmaomega}, we obtain
\be \label{eq:diffusion_plugged_in}
 \left( \frac{3k\phi}{c}i\omega  -   \kappa_n^2 \right) J(n,\sigma,\omega)  &+& \left( \frac{k}{\Delta} \right)^2 \frac{\partial^2J(n,\sigma,\omega)}{\partial\sigma^2} = -\frac{2\kappa_n^2}{R} \frac{\sqrt{6}}{4\pi} \frac{kE}{\Delta^2} \frac{1}{4\pi} \delta(\sigma - \sigma_s).
\ee
At $\sigma=\sigma_s$, continuity must be enforced,
\be \label{eq:matching_condition_1}
J(n, \sigma^-, \omega) = J(n, \sigma^+, \omega),
\ee
and the discontinuity in $dJ/d\sigma$ due to the source is
\be \label{eq:matching_condition_2}
\frac{\partial J(n, \sigma^+, \omega)}{\partial \sigma} - \frac{\partial J(n, \sigma^-, \omega)}{\partial \sigma} & = & 
- \frac{\sqrt{6}}{8} n^2 \frac{E}{kR^3}.
\ee
At large values of $\sigma$ the line profile $\phi$ is small and Eq. \ref{eq:diffusion_plugged_in} becomes
\be \label{eq:diffusion_at_large_sigma}
\frac{\partial^2J}{\partial\sigma^2} \approx \frac{\Delta^2\kappa_n^2}{k^2} J,
\ee
which has solutions 
\be
J(n, \sigma, \omega)\ {\sim}\ e^{\pm \Delta \kappa_n \sigma / k}.
\ee
This approximate solution implies a boundary condition at large $|\sigma|$
\be \label{eq:single_j_derivative}
\frac{\partial J}{\partial \sigma} = \mp \frac{\Delta\kappa_n}{k} J,
\ee
where a negative sign is taken for large $+\sigma$ and a positive sign is taken for large $-\sigma$ to choose the finite solution as $|\sigma|\to 0$. Numerical integrations are performed inward toward $\sigma_s$ over several domains: from large $|\sigma|$ to $\sigma_s$, from large $|\sigma|$ to 0, and from 0 to $\sigma_s$, depending on whether $\sigma_s$ is positive or negative. If $\sigma_s=0$, just two integrations are performed inward from large $|\sigma|$ to 0. Initial values for integration are obtained either by setting $J=1$ and $dJ/d\sigma$ from Equation \ref{eq:single_j_derivative} at large $|\sigma|$ or by matching $J$ and $dJ/d\sigma$ at 0. This gives $J$ and $J'$ on either side of $\sigma_s$, where a prime indicates the derivative $\partial/\partial \sigma$. By enforcing the matching conditions Eqs. \ref{eq:matching_condition_1} and \ref{eq:matching_condition_2}, the eigenfunctions $J(n, \sigma, \omega)$ are obtained over the domain of photon frequencies $\sigma$. Since the solutions are linear in the starting conditions, only two integrations with different starting values are necessary.

We now wish to find an expression for the eigenfunctions $J(n, \sigma, \omega)$ in order to express the specific mean intensity $J(r, \sigma, t)$ as a sum over both spatial and frequency modes. To do so, we employ an ansatz that roughly agrees with Monte Carlo results for $x_s=0$ based on numerical calculations of this result. However, the analysis presented in Appendix \ref{app:wkb} suggests the ansatz is incomplete in the case where $x_s \neq 0$ and the solution is asymmetric about the line center. Let us define the damping rate to be $\gamma \equiv i\omega$, which is real and positive for damped solutions. At the eigenvalues $\gamma=\gamma_{nm}$, the response $J(n, \sigma, \omega)$ is resonant. Near these $\gamma_{nm}$ poles, the eigenfunctions will have the form 
\be \label{eq:jnsigmaomega_approx}
J(n,\sigma,-i\gamma) & \simeq \frac{ J_{nm}(\sigma) }{\gamma_{nm} - \gamma} + C(\gamma, \sigma),
\ee
where $C(\gamma, \sigma)$ varies slowly in $\gamma$. An application of the residue theorem gives %TODO: Problematic, but can this be improved given that we know it doesn't work?
\be
J(r,\sigma,t) & \simeq & j_0(\kappa_n r) J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Summing over all spatial modes $n$ and over all eigenmodes $m$ for a given $n$, we obtain
\be \label{eq:Jrsigmat}
J(r,\sigma,t) & = & \sum_{n=1}^\infty j_0(\kappa_n r)  \sum_{m=1}^{\infty} J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
This ansatz captures the contributions from $n \times m$ simple poles. Taking a derivative with respect to $r$ and evaluating at the surface $r=R$, we use
\be
\frac{dj_0(\kappa_n r)}{dr} \bigg\rvert_R & =& \frac{d}{dr} \left[ \frac{\sin(\kappa_n r)}{\kappa_n r} \right]\bigg\rvert_R
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)
\nonumber \\ & = & \frac{(-1)^n}{R}
\ee
to obtain the flux, which is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Multiplying by $4\pi R^2$ gives the energy per time per frequency emerging from the sphere to be
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\label{eq:dEdtdnu}
\ee
Integrating over time yields a factor $1/\gamma_{nm}$, and by integrating over $d\nu$ we find
\be \label{eq:sum_rule}
E & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3k} \sum_{nm} (-1)^{n+1} \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial ``sum rule'' provides a check on the values of $\gamma_{nm}$ and $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each mode is
\be \label{eq:pnmsoln}
P_{nm} & \equiv & \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^{n+1} \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients are negative for odd values of $n$ and positive for even $n$. The size of each contribution scales roughly as $0.5/(m-7/8)^{2/3}$, with a weak dependence on $n$. This indicates the need for a large number of $n$ and $m$ to converge, in that it takes roughly ten times as many $m$ modes to reduce the size of $P_{nm}$ by a factor of ${\sim}$5.

@@

We seek now to calculate the eigenmodes $J_{nm}(\sigma)$ and eigenfrequencies $\gamma_{nm}$ for a given spatial $n$. These will be labelled by an index $m=1, 2, ...$. To measure the size of the response to detect where resonances occur, we sum the absolute value of $J(n,\sigma,-i\gamma)$ over the array $\sigma$. We call this response $f$, and use the index $j$ to represent the value of the response at discrete points $\gamma_j$ over a range of $\gamma$. In places where $f_j > f_{j-1}$ and $f_j>f_{j+1}$, we have bracketed a resonance that occurs in the interval $(\gamma_{j-1},\gamma_{j+1})$. We then hone in on the eigenvalue before continuing the sweep in $\gamma$. To refine the eigenmode solution, we evaluate $f_{j-1}$, $f_j$, and $f_{j+1}$ at the points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$. Assuming the form in Equation \ref{eq:jnsigmaomega_approx}, a guess at the correct eigenvalue $\gamma_{nm}$ can be calculated from
\be
\gamma_{\rm guess} &=& \frac{b\gamma_{j-1} - \gamma_{j+1}}{b - 1},
\ee
where
\be
b &=& \left(\frac{f_{j} - f_{j-1}}{f_{j} - f_{j+1}}\right)\left(\frac{\gamma_{j+1}-\gamma_{j}}{\gamma_{j-1}-\gamma_{j}}\right).
\ee
This is repeated by replacing initial points $(\gamma_{j-1},\gamma_{j},\gamma_{j+1})$ with closer estimates while the size of the response grows as the resonance is approached. After iterating an eigenvalue $\gamma_{nm}$ to convergence, we now find the corresponding eigenfunction $J_{nm}(\sigma)$. We can solve for $J_{nm}(\sigma)$ by comparing two points $\gamma_1$ and $\gamma_2$ near the resonance, cancelling $C(\gamma, \sigma)$ in the difference to find
\be
J_{nm}(\sigma) & \simeq & \frac{ J(n,\sigma,-i\gamma_1)  - J(n,\sigma,-i\gamma_2) }{ 1/(\gamma_{nm}-\gamma_1) - 1/(\gamma_{nm}-\gamma_2)}.
\ee
\begin{figure}
    \centering
    \includegraphics{Jsoln_n1_m100.pdf}
    \caption{Eigenfunctions $J_{nm}(x)$ for the lowest-order spatial eigenmode $n=1$, and $m=1, ..., 100$ with $x_s=0$, $a = 4.72\times 10^{-4}$ and $\tau_0=10^7$. } %TODO: Mentioned in previous caption that 'a' needs to be included. What's the solution here? Include or not? Reference?
    \label{fig:jsoln}
\end{figure}
\noindent The form of a single eigenmode $J_{nm}(\sigma)$ is oscillatory out to some turning point, $\sigma_{\rm tp}$, at which point the function becomes evanescent. The location of the turning point can be found by ignoring the delta-function discontinuity at the source frequency $\sigma_s$ in Eq. \ref{eq:diffusion_plugged_in} and examining the resulting homogeneous differential equation. We obtain
\be \label{eq:wkb_differential_eqn}
\frac{d^2J}{d\sigma^2} & = & \left[ \left( \frac{\kappa_n \Delta }{k} \right)^2 - \frac{3\phi \gamma\Delta^2}{ck}\right] J,
\ee
where the line profile is approximated as in Eq. \ref{eq:app:line_profile_wing}. When the coefficient on the right hand side is positive, exponential growth or decaying evanescent solutions are found. This occurs on the line wings. When the coefficient on the right hand side is negative, oscillatory solutions are found (propagation), which occurs near the line core. The boundary between propagation and evanescence occurs at the turning point, given by
\be \label{eq:sigma_tp}
\sigma_{\rm tp} & = & \sqrt{\frac{2a}{\pi}}\left( \frac{k \gamma}{ \kappa_n^2 c \Delta} \right)^{3/2}
\ee
Thus, to ensure accuracy in each term of Eq. \ref{eq:Jrsigmat}, the bounds of $\sigma$ must be set sufficiently far outside of $\sigma_{\rm tp}$ such that the function is small at the edges. The scale of an $e$-folding in $J_{nm}(\sigma)$ is $k/(\kappa_n \Delta) = \tau_0 / (\sqrt{\pi} n)$, so a grid of $\sigma$ is chosen that spans a large enough number of $e$-foldings that no oscillatory behavior is present at the boundaries of the domain.

The eigenfunction's oscillatory forms have varying amplitudes which sum to create the final form of the mean intensity. The largest contribution at late times always comes from the ($n=1, m=1$) lowest-order eigenfunction. Fig. \ref{fig:jsoln} shows a set of eigenfunctions $J_{nm}(\sigma)$ to illustrate their relative scales for different $m$ at a fixed spatial eigenmode $n$. The overall scale of the $J_{nm}(\sigma)$ are set by the factor $E/(kR^3)$ with $E$ arbitrarily set to 1. For H atoms with $T=10^4$ K and $\tau_0=10^7$, an eigenfunction has typical size ${\sim} E a / \left(R^2 \Delta \right) = 10^{-37}$ in units of specific mean intensity times time. Additional terms add smaller-magnitude, faster-oscillating components that lead to higher accuracy upon summation with the lower-order terms. The oscillations of various modes must cancel, so many modes $m$ and $n$ are required for convergence to the correct solution. 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{gamma_nm.pdf}
    \caption{Decay rate vs. mode number. Overlapping solid lines are plotted for each $n$. The thickness of the lines decrease with $n$ in order to clearly show where they are in agreement. These numerically-obtained resonant frequencies are compared with the analytic expression in Eq. \ref{eq:gamma_nm}. The parameters for the calculation are $\tau_0=10^7$ and $x_s=0.0$.}
    \label{fig:gamma_nm}
\end{figure}
The values of the $\gamma_{nm}$ can be described approximately with Eq. \ref{eq:gamma_nm}. Their values depend on $m$, $n$, and other physical parameters according to 
\be \label{eq:gamma_nm}
\gamma_{nm} = 2^{-1/3} \pi^{13/6} n^{4/3}\left(m-\frac{7}{8}\right)^{2/3}\frac{c}{R}(a\tau_0)^{-1/3}
\ee
as shown in Appendix \ref{app:wkb}. The power law in $m$ is weak, requiring up to $m=500$ to reduce the scale of $\gamma_{nm}^{\ \ -1}$ by two orders of magnitude. When sweeping through to find resonances, Equation \ref{eq:gamma_nm} is used to set the scale of the sweep points $\gamma_j$ to ensure no $\gamma_{nm}$ are missed. The continuity and close agreement with the analytic expression shown in Fig. \ref{fig:gamma_nm} indicate that the numerical solutions are accurate.

@@

We now calculate the distribution of the times it takes each photon to escape the sphere. This is obtained by integrating Eq. \ref{eq:dEdtdnu} over all frequencies. We find
\be
P(t)  & = & \sqrt{\frac{3}{2}} \frac{16\pi^2 R \Delta^2 }{3kE}     \sum_{nm} (-1)^n  e^{-\gamma_{nm}t} \int d\sigma J_{nm}(\sigma) 
\nonumber \\ & = &  \sum_{nm} P_{nm} \gamma_{nm} e^{-\gamma_{nm}t},
\label{eq:waittime}
\ee
which is normalized to unity. For a sufficiently large number of spatial modes $n$ and frequency modes $m$, the result of this sum can agree with Monte Carlo escape time distributions when $x_s=0$. The late-time distribution is simply an exponential falloff. The rate constant of the falloff is the lowest-order eigenfrequency, $\gamma_{11}$, and its scale is determined by the coefficient $P_{11}$ as in Eq. \ref{eq:pnmsoln}. Thus, an approximate ``fitting function'' that captures both the peak of the escape time distribution and the exponential falloff is
\be \label{eq:fitting_function}
P(t) = \exp{\left[-\left(\frac{t_{\rm diff}}{t}\right)^2\right]} \times \gamma_{11} P_{11} e^{-\gamma_{11}t}.
\ee
The first term represents the early-time distribution, which then transitions to an exponential falloff past a point $ct_{\rm diff}/R = (a\tau_0)^{1/3}$, where $t_{\rm diff}$ is the characteristic diffusion timescale.

In Fig. \ref{fig:tau_scaling}, it is shown that the time constant of exponential decay in fitted Monte Carlo escape time distributions converges with $\gamma_{11}^{-1}$ at sufficiently high $\tau_0$, following a $t\propto(a\tau_0)^{1/3}$ scaling. At lower $\tau_0$, the effects of line core scattering are most important, leading to a larger discrepancy in the characteristic escape timescale. Here, the Monte Carlo accurately includes the photons which scatter in the core many times before escaping, while the semi-analytic solution does not capture this behavior as it uses only the Lorentzian piece of the line profile, and also does not use enough spatial modes to accurately model the frequency regime near line center. However, as $\tau_0$ grows, the effect of core scattering becomes smaller and the approximations hold, agreeing well with the expected $(a\tau_0)^{1/3}$ scaling for the rate constant at late times.
